### 2016-10-11 paper.tex
Intro et bio

%\documentclass[journal,peerreview,onecolumn]{IEEEtran}
\documentclass[journal]{IEEEtran}
\usepackage{etex}
\usepackage{cite}
\usepackage[pdftex]{graphicx}
\graphicspath{{./Fig/}}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage{fixltx2e}
\usepackage{url}
\usepackage{listings}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{tabularx}
\usepackage{empheq}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{minted}\definecolor{bg}{rgb}{0.95,0.95,0.95}
\newcommand\MyBox[2]{
  \fbox{\lower0.75cm
    \vbox to 1.2cm{\vfil
      \hbox to 1.2cm{\hfil\parbox{1.4cm}{#1\\#2}\hfil}
      \vfil}%
  }%
}

% for peerreview
% \usepackage{setspace}
% \onehalfspacing

\newtheorem{prop}{Proposition}

% correct bad hyphenation here
% \hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Operational Feature Selection with Gaussian Mixture Models for remote sensing images}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Adrien~Lagrange,~Mathieu~Fauvel~and~Manuel~Grizonnet% <-this % stops a space
\thanks{A. Lagrange and M. Fauvel are with the Universit\'{e} de Toulouse,
INP-ENSAT, UMR 1201 DYNAFOR, France and with the INRA, UMR 1201
DYNAFOR, France.}%
\thanks{M. Grizonnet is with Centre National d'\'{E}tudes Spatiales, French Space Agency, Toulouse, France.}%
\thanks{This  work was  supported  by the  French National  Research Agency  (ANR)  under  Project Grant  ANR-13-JS02-0005-01  (Asterix project).}}

% The paper headers
\markboth{Transactions on Computational Imaging,~Special Issue on Computational Imaging for Earth Sciences, September~2016}{}

% make the title area
\maketitle

% \tableofcontents
% \clearpage
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
This article presents a forward feature selection algorithm based on Gaussian mixture model (GMM) classifiers. The algorithm selects iteratively features that maximize a criterion function which can be either a classification rate or a measure of divergence. Several variations of this algorithm are explored by changing the criterion function and also by testing a floating forward variation allowing backward step to discard already selected features.

An important effort is made in exploiting GMM properties to implement a fast algorithm. In particular, update rules of the GMM model are used to compute the criterion function with various sets of features. The result is a C++ remote module for the remote sensing processing toolbox Orfeo (OTB) developed by CNES.

Finally, the method is compared to standard classifiers using two different data-sets, one of hyperspectral images with a lot of spectral variables and one with heterogeneous spatial features. The results validate the fact that the method performs well in terms of processing time and classification accuracy in comparison to the standard classifiers available in OTB.\\
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
% \begin{IEEEkeywords}
% remote sensing, hyperspectral imaging, feature selection, gaussian mixture model, fast computing.
% \end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
\label{sec:intro}

\IEEEPARstart{W}{ith} the increasing number of remote sensing missions, the quantity of available data for a given landscape becomes larger and larger. Satellite missions produce a huge amount of data on a regularly (daily) basis. From 2018, the EnMAP (Environmental Mapping and Analysis Program) satellites missioned by the German space agency will produce images with 244 spectral bands with a resolution of 30x30m with a frequency of revisit of 4 days\cite{Müller09enmap}. The Hyperspectral Infrared Imager (HyspIRI) of NASA will also deliver images with 212 spectral bands. Additionally to hyperspectral data, the amount of available hypertemporal data increases a lot. For instance, the European satellites Sentinel-2 were launched successfully recently and 2 Terabytes of data will be released every days~\cite{drusch2012sentinel}. This high volume of earth observation data provides an accurate information of land state and functioning, and help to improve the understanding of the earth~\cite{rs1010001}. However, processing such data is more and more challenging because of statistical and computational issues.

In the spectral or temporal domain, a pixel is represented by a vector
for   wich  each   component   corresponds   to  a   spectral/temporal
measurement.  The  size of  the vector  is the  number of  spectral or
temporal  measurements.   For  hyperspectral images,  this  number  is
typically   about   several   hundreds  while   for   the   Sentinel-2
multitemporal images, the number of spectro-temporal measurement for a
given  year  is approximately  one  thousand.   When working  in  high
dimensional  space,  statistical  methods  made for  low  or  moderate
dimensional  space do  not  adapt  well.  For  instance,  the rate  of
convergence of the statistical estimation decreases when the dimension
grows while conjointly the number of parameters to estimate increases,
making    the    estimation    of   the    model    parameters    very
difficult~\cite{donoho}.  Consequently,  with a limited  training set,
beyond a certain limit, the classification accuracy actually decreases
as the number of features increases~\cite{hughes}.  For the purpose of
classification,  these  problems are  related  to  the \emph{curse  of
  dimensionality}. This is a major drawback in remote sensing since it
is  difficult to  collect a  large and  an accurate  ground-truth.  An
intensive work has  been performed in the remote  sensing community to
build  accurate classifiers  for  high  dimensional images.   Bayesian
models~\cite{book:landgrebe}, feature extraction and feature reduction
techniques~\cite{book:landgrebe,DR:guided:tour},                random
forest~\cite{1396322},   neural  networks~\cite{5411821}   and  kernel
methods~\cite{kernel:methods:rs}  have   been  investigated   for  the
classification of such images.

The  size of  the data  is increasing  with respect  to the  number of
measurement per  pixel. The data  volume of an hyperspectral  image is
typically    several    Gigabytes     per    acquisition    ($\approx$
300km$^2$). Multitemporal data are  now available freely from internet
stream    (see    for    instance     the    Copernicus    data    hub
\url{https://cophub.copernicus.eu/}).  This very  large volume of data
requires   specific   computing  infrastructure.    High   performance
computing   is   actually   investigated   by   the   remote   sensing
community~\cite{christophe2011remote,plaza2011high}.  Main  issues are
related to the use of  parallel approach (multicore, CPU, clusters) to
improve the  processing time, and  to the use of  streaming techniques
when data  does not fit in  memory.  One popular open  source software
solution  is  the Orfeo  Toolbox,  developed  by the  French  National
Agency~\cite{christophe2008orfeo}.   Streaming and  parallel computing
are  convenientely  proposed  to  the  user  through  several  modules
``\emph{ready to use}''.

A method  to reduce  both statistical and  computational issues  is to
perform a reduction of the dimension. In fact, with the \emph{curse of
  dimensionality}     comes     the      \emph{blessing     of     the
  dimensionality}~\cite{bouveyron2014model}:  high   dimensional  data
space   exhibits  interesting   properties   for  the   classification
purpose. In  particular, it is possible  to extract a set  of relevant
features    to   get    a   parsimonious    representation   of    the
data~\cite{jimenez1998supervised}.    For   instance,  in   land-cover
classification,  given  a  set   of  spatial,  temporal  and  spectral
features,  it  is  possible  to  extract  those  which  are  the  most
discriminant      for      the     purpose      of      classification
\cite{fassnacht2014comparison}.   In  hyperspectral   data,  from  the
hundreds of available spectral channels,  it is possible to reduce the
number of channels  to make the processing more efficient  in terms of
statistical complexity  and computational load. In  short, by reducing
the dimension better classification results is expected with a reduced
computational load.

% Among the various dimension reduction techniques, feature selection demonstrates several appealing for practical situation. The objective of feature selection is to find a subset of the original features with a minimum loss according the objective. Thus,

% retrieve the most compact representation of the data with minimum loss of information \cite{Guyon:2006:FEF:1208773}. By deleting the redundant information, feature selection allows to reduce dimension and thus avoid statistical issue, increase algorithm speed and limit storage requirements. Moreover, it adds several beneficial side effects. It improves data understanding by identify where is the most relevant information and it is then possible to save resources when organizing a new data acquisition.

There are two ways to reduce dimension~\cite[Chapter 1]{Guyon:2006:FEF:1208773}: features extraction and feature selection. Feature extraction means reformulate and summarize the information by creating new features in combining the existing ones, it is sometimes referred to \emph{feature construction}. Linear combination of the initial features can be extracted using Principal Component Analysis (PCA)~\cite{jimenez1998supervised} or Independent Component Analysis \cite{villa2011hyperspectral}. Supervised extraction method has also been investigated such as Fischer discriminant analysis and decision boundary feature extraction~\cite{book:landgrebe}. To the contrary, features selection selects a subset of existing features identified as the most relevant for a given criterion. It has the additional advantage to be much more interpretable for the end-user.

Features selection/extraction algorithms can be divided into three classes. The first class is called \emph{filter methods}. They select features independently of the classifier. Features are ranked according to some statistical measures (correlation, independence). For example, the PCA is a typical unsupervised filter method. Bruzzone \emph{et al.}\cite{bruzzone1995extension} develop a supervised filter method based on Jeffries-Matusita distance to maximize the separability of class distribution. The use of correlation between bands has been explored with hyperspectral data\cite{demir2008phase}. In general, these methods have the advantage to be fast and usable with any classifier but they do not take into account the strength and weakness of the chosen classifier and do not optimize directly the classification process.

The second class are known as \emph{wrapper methods}. They search for the best subset of variables for a given learning model. Since exhaustive searches are too expensive in terms of processing time, several sub-optimal search strategies have been designed, mainly iterative forward or backward search~\cite{whitney1971direct,marill1963effectiveness} or a combination of both~\cite{somol1999adaptive}. The advantage of such methods compare to filter methods is that they are dedicated to a particular model and to a particular learning problem. On the other hand, as they require the training of multiple models to test various set of variables, they tend to be more time consuming.

% The method proposed in this paper is actually a wrapper method which has been optimized in order to accelerate model learning and testing which is the weak point of this family.

The third class corresponds to the \emph{embedded method}. They do not separate the features selection process from the learning algorithm and allow interaction between the two processes. The basic example of such method is the decision tree algorithm in which a feature is selected for the creation of each node. Embedded methods also exist for other models, e.g. SVM \cite{guyon2002gene,weston2003use,tuia2015multiclass}.

Despite a large diversity of methods, feature selection algorithms usually do not scale well with the number of pixels to be processed~\cite{fauvel2015fast}. The training computational load is too important to compensate the reduced prediction computational load. Hence, feature selection is not so used in operational situation. However, methods based on Gaussian Mixture Models (GMM) have several interesting properties that make them suitable for feature selection in the context of large amount of data. By taking advantage of their intrinsic properties, it is possible to increase the computational efficiency with respect to standard implementation.

The contribution of the paper is a forward feature selection method using GMM in continuation of~\cite{fauvel2015fast}. An efficient implementation of the method, using linear algebra on block matrix, is presented in order to handle large amount of data. Furthermore, a floating version of the algorithm is proposed. Several criteria are proposed to handle unbalanced training set.  Finally, the developed algorithm is made available as a remote module of the C++ Orfeo Toolbox~\cite{christophe2008orfeo}.

% Finally, tests are conducted to compare different variations of the method and to compare also to other classifiers (GMM, Random Forest, k-nearest-neighbor) available in the Orfeo Toolbox .

The remaining of this article is organized as follows. Section~\ref{sec:gmm-hd} presents GMM classifiers and the problems related to high-dimensional feature space. The features selection methods are detailed in Section~\ref{sec:selection}. The efficient implementation of the proposed selection method is presented in Section~\ref{sec:implementation}. Experimental results on two real high dimensional data set are given Section~\ref{sec:test}. Conclusion and perspectives conclude the paper in Section~\ref{sec:conclusion}.

\section{Gaussian Mixture Models in high dimensional space}
\label{sec:gmm-hd}

The    following    notations    are   used    in    the    remaining.
$\mathcal{S}  = \{\mathbf{x}_i,y_i\}_{i=1}^{n}$  denotes the  training
set where $\mathbf{x}_i \in \mathbb{R}^d$ is the vector of features of
the $i^{th}$ sample, $y_i = 1,...,C$ the associated label, C the total
number of classes,  $n$ the number of samples and  $n_c$ the number of
samples of class $c$.

    \subsection{Gaussian Mixture Models}

    For mixture models, it is assumed  that a given sample $\mathbf{x}$ is
    the realization  of a  random vector which  distribution is  a mixture
    (convex     combination)     of      several     class     conditioned
    distribution~\cite{Fraley00model-basedclustering}:
    \begin{equation}
        p(\mathbf{x}) = \sum_{c=1}^{C} \pi_c f_c(\mathbf{x}|\theta)
    \end{equation}
    where $\pi_c$ is  the prior, \emph{i.e.}, the  proportion of class
    $c$  and  $f_c$  a  parametric density  function  parametrized  by
    $\theta$.

    Among the possible parametric model,  the Gaussian one is the most
    used~\cite{bouveyron2014model}.   It assumes  that each  $f_c$ is,
    conditionally  to  $c$,  a  Gaussian  distribution  of  parameters
    $\boldsymbol{\mu}_c$    and    $\boldsymbol{\Sigma}_c$:
    \begin{equation*}
        f_c(\mathbf{x}|\boldsymbol{\mu}_c, \boldsymbol{\Sigma}_c) = \frac{1}{(2\pi)^{\frac{d}{2}} |\boldsymbol{\Sigma}_c|^{\frac{1}{2}}} \exp \left( -\frac{1}{2} (\mathbf{x} - \boldsymbol{\mu}_c)^t \boldsymbol{\Sigma}_c^{-1} (\mathbf{x} - \boldsymbol{\mu}_c) \right).
    \end{equation*}

    It  is  referred  to  as  Gaussian  mixture  model  (GMM).   In  a
    supervised    learning    framework,    the    class    parameters
    $\boldsymbol{\mu}_c$,   $\boldsymbol{\Sigma}_c$   and  the   prior
    $\pi_c$ are  usually estimated  through the  conventional unbiased
    empirical estimators:
    \begin{align}
        \hat{\pi}_c &= \frac{n_c}{n},\\
        \hat{\boldsymbol{\mu}}_c &= \frac{1}{n_c} \sum_{\{i|y_i = c\}} \mathbf{x}_i ,\\
        \hat{\boldsymbol{\Sigma}}_c &= \frac{1}{(n_c - 1)} \sum_{\{i|y_i = c\}} (\mathbf{x}_i - \boldsymbol{\mu}_c) (\boldsymbol{}x_i - \boldsymbol{\mu}_c)^t.
    \end{align}
    To predict the  class of a new unseen sample,  the maximum \emph{a
      posteriori}  rule  is  used:
    \begin{equation*}
        \mathbf{x} \text{ belongs to } c \Leftrightarrow c = \text{arg} \max_{c \in C} p(c) p(\mathbf{x}|c).
    \end{equation*}
    Under the GMM,  and identifying $p(c)$ as $\pi_c$  and $p(\mathbf{x}|c)$ as
    $f_c(\mathbf{x}|\theta)$ and by taking the log, a decision function is obtained
    \begin{eqnarray}\label{eq:decision}
      Q_c(\mathbf{x}) &=& 2 \log \left( p(c) p(\mathbf{x}|c) \right) \nonumber \\
                      &=& - (\mathbf{x} - \boldsymbol{\mu}_c)^t \boldsymbol{\Sigma}_c^{-1} (\mathbf{x} - \boldsymbol{\mu}_c) \nonumber \\
                      & &-\log (|\boldsymbol{\Sigma}_c|) + 2 \log (\pi_c) - d \log (2\pi).
    \end{eqnarray}

    \subsection{Curse of dimensionality in GMM}
    \label{sec:curse:gmm}

    The computation of eq.~(\ref{eq:decision}) requires the inversion of the
    covariance matrix and the  computation of the logarithm of the
    determinant.  The estimation of these terms suffers from the curse
    of  dimensionality\cite{bouveyron2014model}, because  in practice,
    the  number of  parameters  $\eta_c$ to  estimate  for each  class
    increases quadratically with respect to the number of features, as
    illustrated in Figure~\ref{fig:nb-param}.  Hence, if the number of
    observation  $n_c$ is  small  compared to  the  number of parameters
    $\eta_c$, the  estimated covariance  matrix is badly  conditioned and
    thus the computation  of its inverse and its  determinant would be
    unstable.   The worst  situation  is $n_c<\eta_c$  which leads  to
    a singular covariance matrix.  Unfortunately, this situation happens
    regularly in remote sensing.   For instance in hyperspectral image
    classification,  very few  labeled samples  are usually  available
    because of the difficulty and the high cost to collect ground-truth.

    \begin{figure}[!t]
        \centering
        \begin{tikzpicture}
          \begin{axis}[ymode=log,xmin=0,xmax=200,ymin=0,width=0.8\columnwidth,grid,axis x line=left ,axis y line=left, tick align=outside,xlabel = $d$,ylabel = $\eta_c$,small]
            \addplot+[very thick,mark=none,smooth,domain=0:200,samples=200] (\x,{\x*(\x+3)/2+1});
            \end{axis}
        \end{tikzpicture}
        \caption{Number of parameters $\eta_c$ by class in function of dimension $d$: $\eta_c=d(d+3)/2+1$.\label{fig:nb-param}}
    \end{figure}

    There are two major solutions to this problem. First option is to stabilize the inversion of the covariance matrices. Some methods investigate the use of constraints on the direct problem. Reynolds \emph{et al.} \cite{reynolds1995robust} proposed to use diagonal covariance matrices. It is also possible to force the diagonal element to be higher than a given value by maximizing the GMM likelihood \cite{hathaway1985constrained}. Celeux and Govaert \cite{celeux1995gaussian} suggest to use equality constraints between coefficients of the covariance matrix in a parsimonious cluster-based GMM framework. Others papers propose to work on the inverse problem. A classical method is to use a regularization method as the well-known ridge regularization \cite{hoerl1970ridge}. A ridge regularization aims to stabilize the inversion by replacing the covariance matrix $\boldsymbol{\Sigma}_c$ by $\boldsymbol{\Sigma}_c + \tau \mathbf{I}$ where $\tau$ is a positive parameter and $\mathbf{I}$ the identity matrix. Jensen \emph{et al.} \cite{jensen2008regression} propose a different approach using a sparsity approximation to inverse the covariance matrix.

    %% Question: tu suggérais de reprendre ici "Ensuite, il y a la réduction de dimension données des refs (voir intro du papier JSTARS) et dans la réduction  de dimension, il  y a  la sélection de features." Le soucis que je vois, c'est que je détaille déjà dans l'intro les différentes méthodes de sélection/extraction et je trouve que ça devient pas très clair si je reprends tout en me basant sur le papier JSTARS. J'aurais dit avt "il y a trois types de sélection: filtre, wrapper, embedded blablabla" et ensuite ça me fait reprendre ici une autre distinction "supervisé/non-supervisé" si je me base sur jstars qui est un peu redondante il me semble. Qu'en dis-tu ?
    Second option is to reduce dimension. Features extraction/selection methods have been developed in order to reduce the dimension with various approaches described in Section~\ref{sec:intro}. In this study, this latter option is explored and a feature selection method named sequential forward features selection is presented after a brief overview of the selection algorithms in the next Section.

\section{Sequential forward features selection}
\label{sec:selection}

The feature selection method proposed in this work is a wrapper method associated to GMM models. Two elements are needed to set up a wrapper method:
\begin{enumerate}
\item A function that  ranks the features according to some good classification criteria,
\item A search strategy to optimize the function.
\end{enumerate}

Section~\ref{sec:criterion} describes the different criterion used in this work and two search strategies are discussed in~\ref{sec:selection:method}.

    \subsection{Criterion function}
    \label{sec:criterion}
    The criterion evaluates how a given model built with a subset of features performs for the classification task. It can be an estimation of the correct classification or a measure of separability/similarity between class distribution. The former are in general more demanding in terms of processing time than the later.

    % Criterion functions aim to evaluate either a rate of correct classification or the separability/similarity of class distributions. These functions are used to estimate which sets of variables are the best to represent data, to assure the separability of the classes and to perform classification. The choice of a criterion function is the choice of a way to compare sets of variables.

        \subsubsection{Measures of correct classification}
        \label{sec:criterion-rate}

        A measure of correct classification is based on an error matrix $M$, or \emph{confusion matrix}~\cite[Chapter 4]{congalton2008assessing}. An element $M_{ij}$ of the matrix represents the number of samples of class $i$ classified as class $j$. The confusion matrix allows the computation of several global and per-class indices related to the classification accuracy~\cite{congalton2008assessing}. Three global criteria were used :

        \begin{itemize}
        \item  \emph{The overall  accuracy} (OA)  is the  rate of  the
          number of samples with the  correct predicted label over the
          total  number of  samples~\cite{congalton2008assessing}. This
          metric is  easy to interpret  but is  biased in the  case of
          unbalanced classes.

        \item \emph{The Cohen's kappa} (K) is a statistic which measures the probability of agreement between predictions and ground-truth~\cite{congalton2008assessing}.

        \item \emph{The mean F1 score}  (F1mean) is the average of the
          F1 score  for each class  and the  F1 score is  the harmonic
          mean of  the precision  (number of  True Positive  over True
          Positive plus False Positive) and the recall (number of True
          Positive over True Positive plus False Negative)~\cite{powers2011evaluation}.
        \end{itemize}
        High  value  of  theses  indices  correspond  to  an  accurate
        classification.

        The classification rate is  estimated by a $n_{cv}$- cross-validation ($n_{cv}$-CV)~\cite{opac-b1127878}.  To compute the $n_{cv}$-CV, a subset is removed  from $\mathcal{S}$ and  the GMM is learned with the remaining training samples. A test error is computed with the removed  training  samples  used  as  validation  samples.   The process is  iterated $n_{cv}$ times and  the estimated classification rate is computed as the mean  test error over the $n_{cv}$ subsets of $\mathcal{S}$.

        \hspace{0pt} \\

        \subsubsection{Similarity between distributions}
        The similarity  between two distributions is  quantified using
        divergence  measure~\cite{opac-b1097517}. Contrary  to measure
        of correct  classification, divergences are  computed directly
        on the trained model, with no need of validation samples.  Two
        particular   divergences   are   used  in   this   work:   the
        \emph{Kullback-Leibler}        divergence       and        the
        \emph{Jeffries-Matusita}  distance.  The  advantage  of  these
        divergences is  that they have  an explicit expression  in the
        case of Gaussian models. The  simplification allows to get rid
        of  any integration  calculus which  is a  major problem  when
        dealing with high-dimensional data.

        \emph{The   Kullback-Leibler   divergence}   (KL   divergence)
        measures  the  amount  of  information  lost  when  the  first
        distribution     is     approximated     by     the     second
        one\cite{kullback1987letter}. It can be explicitly computed in
        the case of Gaussian distributions:
        \begin{align}
            &KL_{ij} = \frac{1}{2} \biggl\{ \text{Tr} (\boldsymbol{\Sigma}_i^{-1} \boldsymbol{\Sigma}_j) \nonumber \\
            & + (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t \boldsymbol{\Sigma}_i^{-1} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) - d + \log \left( \frac{|\boldsymbol{\Sigma}_i|}{|\boldsymbol{\Sigma}_j|} \right) \biggr\},
        \end{align}
        where Tr is the trace operator and $d$ the dimension of the distribution.

        The    KL   divergence    is   not    symmetric,   \emph{i.e.}
        $KL_{ij} \ne KL_{ji}$.  A symmetrized version  is used to
        compute the criterion function:
        \begin{align}\label{eq:skl}
          SKL_{ij} &=KL_{ij} + KL_{i,j} \nonumber \\
            &= \frac{1}{2} \biggl\{ \text{Tr} (\boldsymbol{\Sigma}_i^{-1} \boldsymbol{\Sigma}_j + \boldsymbol{\Sigma}_j^{-1} \boldsymbol{\Sigma}_i) \nonumber \\
            &~~+ (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t (\boldsymbol{\Sigma}_i^{-1} + \boldsymbol{\Sigma}_j^{-1}) (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) - 2d \biggr\}.
        \end{align}

        The extension  to the multiclass  situation is done  by taking
        the weighted mean  of the KL divergences computed  on all pair
        of classes~\cite{bruzzone1995extension}:
        \begin{equation}
            C_{SKL} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_i \pi_j SKL_{ij}.
        \end{equation}

        \emph{The Bhattacharyya distance} is defined in the case of Gaussian model as
        \begin{align}
            {B}_{ij} = &\frac{1}{8} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t \left( \frac{\boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j}{2} \right)^{-1} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) \nonumber \\
            &+ \frac{1}{2} \ln \left( \frac{|\boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j|}{\sqrt{|\boldsymbol{\Sigma}_i| |\boldsymbol{\Sigma}_j|}} \right).
        \end{align}

        The \emph{Jeffries–Matusita distance} is a measure based on the Bhattacharyya distance. It saturates if the separability between the two distribution increases~\cite{bruzzone2009novel}. The JM distance is defined as
        \begin{equation}\label{eq:jm}
            {JM}_{ij} = \sqrt{ 2 \{1 - \text{exp}(-B_{ij})\} }.
        \end{equation}

        Similar to the KL divergence,  a weighted mean of the distance
        between two classes is computed to aggregate the measures in a
        single value:
        \begin{equation}
            {C}_{JM} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_i \pi_j {JM}_{ij}.
        \end{equation}

        \vspace{10 mm}

        % According to \cite{bruzzone2009novel}, it is interesting to notice that the KL divergence increases quadratically with respect to the distance between the mean vectors of the class distributions whereas the measures of correct classification asymptotically tends to one when distributions are perfectly separable. On the contrary, the JM distance tends to saturate as the measures of correct classification.

        Table~\ref{tab:crit} summarizes the presented criterion functions and their characteristics.

        \begin{table}[!t]
            \centering
            \caption{Summary of the different criterion functions.\label{tab:crit}}
            \begin{tabular}[b]{lcc}
              \toprule
              Criterion & Type & Complexity \\
              \midrule
              Overall accuracy            & Accuracy   & High \\
              Cohen's kappa               & Accuracy   & High\\
              F1 mean                     & Accuracy   & High\\
              \midrule
              Kullback-Leibler divergences & Divergence & Low \\
              Jeffries-Matusita distance  & Divergence & Low \\
              \bottomrule
            \end{tabular}
        \end{table}

        \subsection{Selection method}
        \label{sec:selection:method}
        Two  sequential  search  algorithms were  implemented  in  this
        work~\cite{Guyon:2006:FEF:1208773}:  the   sequential  forward
        selection and the sequential  floating forward.  The later one
        is  an   extension  of   the  former.  Both   select  features
        iteratively.

        \subsubsection{Sequential forward features selection}
        \label{sec:forward-presentation}

        The Sequential Forward Selection (SFS) starts with an empty set of selected features.  At each  step, the feature that most  improves the criterion function $J$ selected in Table~\ref{tab:crit} is added  to the set. This feature is definitively added to the pool of selected features and the algorithm stops when a given number of variables \emph{maxVarNb} has been selected. The Algorithm~\ref{alg:sfs} presents the process in details.

        \begin{algorithm}
        \caption{Sequential forward features selection\label{alg:sfs}}
        {\footnotesize
        \begin{algorithmic}[1]
        \REQUIRE $\Omega,J,\text{maxVarNb}$
        \STATE $\Omega=\emptyset$
        \STATE $F=\text{\{all variables\}}$
        \WHILE{$\text{card}(\Omega) \leq maxVarNb$}
        \FORALL{$f_i \in F$}
        \STATE $R_i = J(\{\Omega + f_i\})$
        \ENDFOR
        \STATE $j=\text{arg} \max_{i} R_i$
        \STATE $\Omega = \{\Omega + f_j\}$
        \STATE $F = F \setminus f_j$
        \ENDWHILE
        \RETURN $\Omega$
        \end{algorithmic}
        }
        \end{algorithm}

        \subsubsection{Sequential floating forward feature selection}
        \label{sec:floating-presentation}

        The Sequential Floating Forward Selection (SFFS)\cite{somol1999adaptive} is based on two algorithms: the SFS described above and the Sequential Backward Selection (SBS). The SBS is the backward equivalent of SFS. The difference is that it starts with every features in the pool of selected features and tries at each step to remove the less significant one in term of the given criterion function.

        The SFFS works like the SFS but between each step of the SFS algorithm, a backward selection is operated to identify the less important feature. If the criterion value is higher than the best value ever obtained with a set of same size, the identified feature is picked out. The SBS step is repeated while removing the less important feature leads to an increase of the criterion value. Then SFS is called again. The algorithm stops when a given number of features \emph{maxVarNb} has been selected. The Algorithm~\ref{alg:sffs} provides details of the process.

        This SFFS algorithm eventually tests more solutions than the SFS algorithm. The results are expected to be better but the trade-off is an increased computational time which is dependent on the complexity of the dataset.

        \begin{algorithm}
        \caption{Sequential floating forward features selection\label{alg:sffs}}
        {\footnotesize
        \begin{algorithmic}[1]
        \REQUIRE $J,\text{maxVarNb}$
        \STATE $\Omega=\overbrace{(\emptyset,...,\emptyset)}^{maxVarNb}$
        \STATE $F=\text{\{all variables\}}$
        \STATE $k=0$
        \WHILE{$k \neq \text{maxVarNb}$}
        \FORALL{$f_i \in F$}
        \STATE $R_i = J(\{\Omega_k + f_i\})$
        \ENDFOR
        \STATE $j=\text{arg} \max_{i} R_i$
        \STATE $k=k+1$
        \IF{$R_j \geq J(\Omega_k)$}
        \STATE $\Omega_k = \{\Omega_{k-1} + f_j\}$
        \STATE $\text{flag}=1$
        \WHILE{$k > 2 \text{ and } \text{flag}=1$}
        \FORALL{$f_i \in \Omega_k$}
        \STATE $R_i = J(\{\Omega_k \setminus f_i\})$
        \ENDFOR
        \STATE $j=\text{arg} \max_{i} R_i$
        \IF{$R_j > J(\Omega_{k-1})$}
        \STATE $\Omega_{k-1} = \{\Omega_k \setminus f_j\}$
        \STATE $k=k-1$
        \ELSE
        \STATE $\text{flag}=0$
        \ENDIF
        \ENDWHILE
        \ENDIF
        \ENDWHILE
        \RETURN $\Omega_{\text{maxVarNb}}$
        \end{algorithmic}
        }
        \end{algorithm}

\section{Efficient implementation}
\label{sec:implementation}
The most  demanding part  of the  algorithm is  the evaluation  of the
criterion  for  all   the  remaining  variables  (see   lines  5-7  in
Algorithm~\ref{alg:sfs}). Calculus are based on linear algebra, and the
numerical  complexity is  on average  $O(d^3)$.  Furthermore,  for the
\emph{accuracy}-type  criterion the  complexity  is  augmented by  the
cross-validation procedure.

An efficient implementation of the criterion optimization is proposed.
It based on  the symmetry property of the covariance  matrix and block
inverse  formula~\cite{IMM2012-03274}. In  the following  it is  shown
that the criterion  can be split into  two part: one that  needs to be
computed for each  tested variable, and one that needs  to be computed
only once per  selection step. For the  cross-validation part, updates
rules are given to derive sub-models  without the necessity to learn a
GMM models for each fold.

\subsection{Statistical update rules}
        \subsubsection{Update for cross validation}
        \label{sec:update-cv}
        Based on \cite{fauvel2015fast}, a method to accelerate the $n_{cv}$-fold cross-validation process in the case of criterion functions based on correct classification measures was implemented. The idea is to estimate the GMM with the whole training set once and then, instead of training models on $(n_{cv}-1)$ folds, parameters of the complete model are used to derive those of sub-models.

        A sub-GMM model can be estimated at a reduced complexity when one fold is remove from the training set. From~\cite{fauvel2015fast}, the following formulae for unbiased estimators can be derived:
        \begin{prop}[Cross-validation mean update]
            \label{eq:update-cv1}
            \begin{equation*}
                \boldsymbol{\hat{\mu}}_c^{n_c-\nu_c} = \frac{n_c \boldsymbol{\hat{\mu}}_c^{n_c} - \nu_c \boldsymbol{\hat{\mu}}_c^{\nu_c}}{n_c - \nu_c} \nonumber
            \end{equation*}
        \end{prop}
        \begin{prop}[Cross-validation covariance matrix update]
            \label{eq:update-cv2}
            \begin{align*}
              \boldsymbol{\hat{\Sigma}}_c^{n_c-\nu_c} = &\frac{1}{n_c-\nu_c-1} \biggl\{ (n_c-1) \boldsymbol{\hat{\Sigma}}_c^{n_c} - (\nu_c-1)\boldsymbol{\hat{\Sigma}}_c^{\nu_c} \nonumber \\
                                                        &- \frac{n_c \nu_c}{(n_c-\nu_c)} (\boldsymbol{\hat{\mu}}_c^{\nu_c}-\boldsymbol{\hat{\mu}}_c^{n_c})(\boldsymbol{\hat{\mu}}_c^{\nu_c}-\boldsymbol{\hat{\mu}}_c^{n_c})^t \biggr\} \nonumber
            \end{align*}
        \end{prop}
        \noindent where $n_c$ is the number of samples of class $c$, $\nu_c$ is the number of samples of class $c$ removed from the initial set, exponents on $\boldsymbol{\Sigma}_c$ and $\boldsymbol{\nu}_c$ denotes the set of samples used to compute them.

        \subsubsection{Criterion function computation}
        \label{sec:update-crit}
        At iteration $k$, depending on the criterion, three terms have
        to  be computed:  the inverse  of the  covariance matrix,  the
        logarithm of the determinant of  the covariance matrix and the
        quadratic  term in  eq.(~\ref{eq:decision}). However,  all these
        terms  have already  been computed  for iteration  $(k-1)$. By
        using  the  fact that  the  covariance  matrix is  a  positive
        definite     symmetric    matrix     and    block     matrices
        formulae~\cite[Chapter   9.2]{webb2003statistical},    it   is
        possible to factorize these terms at iteration $k$.

        In the  remaining of the  paper, $\boldsymbol{\Sigma}^{(k-1)}$
        denotes the  covariance matrix of the  $(k-1)^{th}$ iteration,
        \emph{i.e.}, the  covariance matrix  of the  selected features
        and $\boldsymbol{\Sigma}^{(k)}$ denotes a covariance matrix at
        the $k^{th}$ iteration, \emph{i.e.},  the covariance matrix of
        an augmented set. Then, since $\boldsymbol{\Sigma}^{(k)}$ is a
        positive definite symmetric matrix,  the covariance matrix can
        be written as

        % Then, the inverse of  the covariance matrix
        % $(\boldsymbol{\Sigma}^{(k)})^{-1}$,   the   quadratical   term
        % $(\mathbf{x}^{(k)})^t         (\boldsymbol{\Sigma}^{(k)})^{-1}
        % \mathbf{x}^{(k)}$        and        the        log-determinant
        % $\log   |\boldsymbol{\Sigma}^{(k)}|$  can   be  expressed   in
        % function of terms of the $(k-1)^{th}$ iteration.

        \begin{equation}\label{eq:cov}
            \boldsymbol{\Sigma}^{(k)} =
            \bigg[\begin{array}{cc}
            \boldsymbol{\Sigma}^{(k-1)} & \mathbf{u}      \\
            \mathbf{u}^t          & \sigma_{kk} \\
            \end{array}\bigg],
        \end{equation}
        where  $\mathbf{u}$  is  the  $k^{th}$ column  of  the  matrix
        without      the      diagonal      element,      \emph{i.e.},
        $\mathbf{u}_{i}   =    \boldsymbol{\Sigma}^{(k)}_{i,k}$   with
        $i  \in [1,k-1]$.  Using  block matrix  inverse formulae,  the
        following proposition holds:
        \begin{prop}[Forward update rule for the inverse of the covariance matrix]
        \label{eq:update-inv}
        \begin{equation}\label{eq:cov:inv}
                (\boldsymbol{\Sigma}^{(k)})^{-1} =
                \bigg[\begin{array}{cc}
                \mathbf{A} & \mathbf{v} \\
                \mathbf{v}^t  & \frac{1}{\alpha} \\
                \end{array}\bigg]
            \end{equation}
        \end{prop}
        \noindent                                                where
        $\mathbf{A}     =     (\boldsymbol{\Sigma}^{(k-1)})^{-1}     +
        \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}
        \mathbf{u}^t              (\boldsymbol{\Sigma}^{(k-1)})^{-1}$,
        $\mathbf{v}           =           -           \frac{1}{\alpha}
        (\boldsymbol{\Sigma}^{(k-1)})^{-1}       \mathbf{u}$       and
        $      \alpha      =      \sigma_{kk}      -      \mathbf{u}^t
        (\boldsymbol{\Sigma}^{(k-1)})^{-1}  \mathbf{u}$.  This  update
        formulae      is     used      to      obtain     all      the
        $(\boldsymbol{\Sigma}^{(k)})^{-1}$  corresponding  to all  the
        possible augmented set of a given selection iteration. Similar
        update formulae can be written for the backward step.

        % In the case of a backward step in SFFS algorithm, the formula is inverted in order to compute $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ knowing $(\boldsymbol{\Sigma}^{(k)})^{-1}$. The update rule becomes (proof in Appendix~\ref{app:proof-update})
        % \begin{prop}[Backward update rule for inverse of covariance matrix]
        % \label{eq:update-inv-back}
        %     \begin{equation*}
        %         (\boldsymbol{\Sigma}^{(k-1)})^{-1} = \underbrace{\mathbf{A}}_{\substack{\text{computed once}\\ \text{per selection step}}} - \underbrace{\alpha \mathbf{v} \mathbf{v}^t}_{\substack{\text{computed for} \\ \text{each augmented set}}}
        %     \end{equation*}
        % \end{prop}

        Using eq.~(\ref{eq:cov}) and~(\ref{eq:cov:inv}), it is possible to
        deduce  the   following  propositions  (proof  are   given  in
        Appendix~\ref{app:proof-update}).
        \begin{prop}[Update rule for the quadratical term]
        \label{eq:update-quad}
            \begin{align}
              (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\underbrace{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}}_{\substack{\text{computed once per selection step}}} \nonumber \\
                                                                                       &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2}_{\substack{\text{computed for each augmented set}}}.
            \end{align}
        \end{prop}
        \begin{prop}[Update rule for logdet]
        \label{eq:update-log}
            \begin{equation}
                \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)}_{\substack{\text{computed once}\\ \text{per selection step}}} + \underbrace{\log \alpha}_{\substack{\text{computed for} \\ \text{each augmented set}}}.
            \end{equation}
        \end{prop}

        From these update rules, it is now possible to split each criterion into two parts: one computed once per selection step and one computed for each augmented set.

        \begin{prop}[Decision function~(\ref{eq:decision})]
            \begin{align}
            \label{eq:q-update}
                Q(\mathbf{x}) = &- \underbrace{(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
                &- \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right) + 2 \log (\pi_c) + k \log(2 \pi)}_{\substack{\text{computed once per selection step}}} \nonumber \\
                &- \underbrace{\alpha \left( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] (\mathbf{x}^{(k)} - \boldsymbol{\mu}^{(k)}) \right)^2 - \log \alpha}_{\substack{\text{computed for each augmented set}}}.
            \end{align}
        \end{prop}

      \begin{prop}[Kullback-Leibler divergence~(\ref{eq:skl})]
        \begin{align}
        \label{eq:skl-update}
            & SKL_{ij} = \frac{1}{2} \biggl\{ \underbrace{\text{\normalfont Tr} \left((\boldsymbol{\Sigma}_i^{(k)})^{-1} \boldsymbol{\Sigma}_j^{(k)} + (\boldsymbol{\Sigma}_j^{(k)})^{-1} \boldsymbol{\Sigma}_i^{(k)}\right)}_{\substack{\text{computed for each augmented set}}} \nonumber \\
            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_i}^t & \frac{1}{\alpha_i} \end{array}\right] (\boldsymbol{\mu}_i^{(k)} - \boldsymbol{\mu}_j^{(k)}) )^2}_{\substack{\text{computed for each augmented set}}} \nonumber \\
            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_j}^t & \frac{1}{\alpha_j} \end{array}\right] (\boldsymbol{\mu}_i^{(k)} - \boldsymbol{\mu}_j^{(k)}) )^2 - 2k}_{\substack{\text{computed for each augmented set}}} \nonumber \\
            &+ \underbrace{(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t (\boldsymbol{\Sigma}_i^{(k-1)})^{-1}(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber\\
          & + \underbrace{(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t(\boldsymbol{\Sigma}_j^{(k-1)})^{-1} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once per selection step}}} \biggr\}.
        \end{align}
        with $(\boldsymbol{\Sigma}_i^{(k)})^{-1}$ computed with Proposition~\ref{eq:update-inv};
      \end{prop}

      \begin{prop}[Jeffries-Matusita distance~(\ref{eq:jm})]
      \begin{align}
        \label{eq:jm-update}
            \text{B}_{ij} = &\underbrace{\frac{1}{4} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{2} \ln \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_i^{(k-1)}| |\boldsymbol{\Sigma}_j^{(k-1)}|}} \right)}_{\substack{\text{computed once per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_{i}^{(k)} - \boldsymbol{\mu}_{j}^{(k)}) )^2 + \frac{1}{2} \ln \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_i \alpha_j}} \right)}_{\substack{\text{computed for each augmented set}}},
      \end{align}
      where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j$ and $\tilde{\alpha}$ and $\mathbf{\tilde{v}}$ are defined as $\alpha$ and $\mathbf{v}$ but using $\boldsymbol{\tilde{\Sigma}}$ instead of $\boldsymbol{\Sigma}$.
    \end{prop}

    The  Algorithm~\ref{alg:sffs-update} illustrates  the optimization
    of the Algorithm~\ref{alg:sffs} using these formulae.

    \subsection{Numerical issues}
    For each  iteration $k$, after  the selection of  optimal features
    w.r.t  the  selected criterion,  the  inverses  of the  covariance
    matrices and their log-determinant  needs to be computed. However,
    the lack of training samples or the highly correlated features may
    induce  a  badly-conditioned  matrix  with  very  small,  or  even
    negative, eigenvalues.   Such values will degrade  drastically the
    estimation of the  inverse and of the log-determinant,  and so the
    numerical stability.

    To deal with this limitation, the choice has been made to perform an eigendecomposition of the covariance matrix $\boldsymbol{\Sigma}^{(k)}$:

    \begin{eqnarray}
    \boldsymbol{\Sigma}^{(k)} = \mathbf{Q}^{(k)} \boldsymbol{\Lambda}^{(k)} (\mathbf{Q}^{(k)})^t\label{eq:eigendecomp}
    \end{eqnarray}

    where  $\boldsymbol{\Lambda}^{(k)}$ and  $\mathbf{Q}^{(k)}$ are  the
    diagonal matrix  of eigenvalues of  the covariance matrix  and the
    orthonormal matrix of corresponding eigenvectors, respectively. To
    prevent numerical instability, the  eigenvalues are thresholded to
    EPS\_FLT. In our  implementation, EPS\_FLT is set  to the floating
    machine precision.

    Then, the inverse of the covariance matrix can be computed as

    \begin{eqnarray}
      \boldsymbol{\Sigma}^{(k)} = \mathbf{Q}^{(k)} (\tilde{\boldsymbol{\Lambda}}^{(k)})^{-1} (\mathbf{Q}^{(k)})^t\label{eq:eigendecomp:inv}
    \end{eqnarray}

    and the log-determinant as

    \begin{eqnarray}
      \label{eq:log:det}
      \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \sum_{i=1}^d\log(\tilde{\lambda}_i^{(k)}),
    \end{eqnarray}
    where the \textasciitilde{} indicated thresholded values.

    Same reasoning applied  to the term $\alpha$ in  the update rules:
    it       is       also        thresholded       to       EPS\_FLT.
    Algorithm~\ref{alg:sffs-update}    details   when    computational
    stability is enforced (lines 6, 10, 20 and 24).

    \begin{algorithm}
    \caption{Sequential floating forward features selection with updates\label{alg:sffs-update}}
    {\footnotesize
    \begin{algorithmic}[1]
    \REQUIRE $J,\text{maxVarNb}$
    \STATE $\Omega=\overbrace{(\emptyset,...,\emptyset)}^{maxVarNb}$
    \STATE $F=\text{\{all variables\}}$
    \STATE $k=0$
    \WHILE{$k \neq \text{maxVarNb}$}
    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)} = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$
    \STATE {\bfseries for all} $\Lambda_{i,i}$ {\bfseries do} $\Lambda_{i,i} = \max (\text{EPS\_FLT},\Lambda_{i,i})$
    \STATE Precompute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions (\ref{eq:update-inv}), (\ref{eq:update-quad}) and (\ref{eq:update-log})
    \FORALL{$f_i \in F$}
    \STATE Compute update constant $\alpha$
    \STATE $\alpha = \max (\text{EPS\_FLT},\alpha)$
    \STATE $R_i = J(\{\Omega_k + f_i\})$ using Equations (\ref{eq:q-update}), (\ref{eq:skl-update}) or (\ref{eq:jm-update})
    \ENDFOR
    \STATE $j=\text{arg} \max_{i} R_i$
    \STATE $k=k+1$
    \IF{$R_j \geq J(\Omega_k)$}
    \STATE $\Omega_k = \{\Omega_{k-1} + f_j\}$
    \STATE $\text{flag}=1$
    \WHILE{$k > 2 \text{ and } \text{flag}=1$}
    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)} = = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$
    \STATE {\bfseries for all} $\Lambda_{i,i}$ {\bfseries do} $\Lambda_{i,i} = \max (\text{EPS\_FLT},\Lambda_{i,i})$
    \STATE Compute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions (\ref{eq:update-inv}), (\ref{eq:update-quad}) and (\ref{eq:update-log})
    \FORALL{$f_i \in \Omega_k$}
    \STATE Compute update constant $\alpha$
    \STATE $\alpha = \max (\text{EPS\_FLT},\alpha)$
    \STATE $R_i = J(\{\Omega_k \setminus f_i\})$ using Equations (\ref{eq:q-update}), (\ref{eq:skl-update}) or (\ref{eq:jm-update})
    \ENDFOR
    \STATE $j=\text{arg} \max_{i} R_i$
    \IF{$R_j > J(\Omega_{k-1})$}
    \STATE $\Omega_{k-1} = \{\Omega_k \setminus f_j\}$
    \STATE $k=k-1$
    \ELSE
    \STATE $\text{flag}=0$
    \ENDIF
    \ENDWHILE
    \ENDIF
    \ENDWHILE
    \RETURN $\Omega_{\text{maxVarNb}}$
    \end{algorithmic}
    }
    \end{algorithm}

    \subsection{Implementation}
    \label{sec:otb-module}

    The proposed  method has been implemented  in \texttt{C++} through
    the  Orfeo  Toolbox (OTB)~\cite{christophe2008orfeo}.   The  Orfeo
    Toolbox  is  an  open-source  library  for  remote  sensing  image
    processing,  developed by  the  French Space  Agency (CNES).   The
    feature selection algorithm  can be installed as  a remote module,
    the           source            code           is           freely
    available\footnote{\url{https://github.com/Laadr/otbFFSforGMM}}.

    Following OTB framework, two application are available. The first,
    called  \texttt{otbcli\_TrainGMMSelectionApp}, performs  a feature
    selection algorithm (SFS or SFFS) with the one given criterion (to
    choose  between  Jeffries-Matusita  dist.,  Kullback-Lieber  div.,
    Cohen’s  kappa, overall  accuracy,  mean  F1-score).  The  second,
    called \texttt{otbcli\_PredictGMMApp}, generates the thematic maps
    according to the  learn model. The only  technical limitation that
    the  training set  must  fit  in the  RAM  of  the computer.   The
    classification step is streamed and there is no limitation in term
    of image  size. The Figure~\ref{fig:otb:ffs} shows  a code excerpt
    to run the application.

    \begin{figure} 
        \centering %Pas touche l'indentation !
        \begin{minted}[fontsize=\tiny,bgcolor=bg]{sh} 
otbcli_TrainGMMSelectionApp -io.il hyper.tif \
                            -io.vd reference.shp \
                            -gmm.varnb 20 -gmm.method forward -gmm.crit jm\
                            -gmm.best 1 -gmm.seed 0\
                            -io.out model.txt

otbcli_PredictGMMApp -in hyper.tif \
                     -model model.txt -modeltype selection\
                     -out ThematicMap.tif
        \end{minted}
        \caption{OTB Module: The feature selection is done on the image \emph{hyper.tif} using the training set from \emph{reference.shp}. The feature selection algorithm is the forward search used with the Jeffries-Matusita criterion, 20 features are extracted and the corresponding GMM model is saved in \emph{model.txt}. Then the whole image is classified using the model and the results is saved in the geotiff \emph{ThematicMap.tif}.}
        \label{fig:otb:ffs}
    \end{figure}

\section{Data-sets}
\label{sec:datasets}
Numerical experiments  have been conducted on  two different data-sets.
The first  dataset, called  \emph{Aisa}, is an  airborne hyperspectral
dataset  and  the  second,  called \emph{Potsdam},  is  an  very  high
resolution multispectral airborne image.

    \subsection{Aisa dataset}
    \label{sec:aisa-dataset}
    The Aisa dataset has been acquired by the AISA Eagle sensor during
    a  flight campaign  over Heves,  Hungary.  It  contains 252  bands
    ranging from  395 to 975  nm. 16 classes  have been defined  for a
    total of 361,971  referenced pixels, Table~\ref{tab:aisa} presents
    the number of pixel per  class.  The Figure~\ref{fig:aisa} shows a
    colored composition of the image and the ground-truth.

    \begin{figure}[!t]
        \centering
        \begin{tabular}{c}
            \includegraphics[width=0.7\columnwidth]{Fig/aisa.png} \\
            {\bfseries{(a)}} \\
            \includegraphics[width=0.7\columnwidth]{Fig/aisa_gt.png} \\
            {\bfseries{(b)}} \\
        \end{tabular}
        \caption{Aisa dataset: {\bfseries{(a)}} colored composition of the image (R: 634nm, G: 519nm, B: 477nm), {\bfseries{(b)}} ground-truth.\label{fig:aisa}}
    \end{figure}

    \begin{table}[!t]
        \centering
        \caption{Information classes for the Aisa dataset.\label{tab:aisa}}
        \begin{tabular}[b]{lc}\toprule
          Class & Number of samples \\ \midrule
          Winter wheat & 136,524 \\
          Sunflower & 61,517 \\
          Green fallow last year treatment & 30,197 \\
          Alfalfa & 17,626 \\
          Maize & 18,278 \\
          Millet & 7,199 \\
          Broadleaved forest & 10,746 \\
          Meadow & 23,283 \\
          Winter barley & 2,799 \\
          Reed & 4,222 \\
          Water course & 4,773 \\
          Rape & 26,566 \\
          Green fallow with shrub & 9,272 \\
          Green fallow last year treated & 3,426 \\
          Pasture & 2,107 \\
          Oat & 3,436 \\ \bottomrule
        \end{tabular}
    \end{table}

    \subsection{Potsdam dataset}
    \label{sec:pots-dataset}

    This  second dataset  is built  from a  dataset of  remote sensing
    images distributed by the International Society for Photogrammetry
    and                         Remote                         Sensing
    (ISPRS)\footnote{\url{http://www2.isprs.org/commissions/comm3/wg4/2d-sem-label-potsdam.html}}. The
    dataset is composed of aerial images of the urban area of Potsdam.
    The area  is divided  into 38  patches of  6000$\times$6000 pixels
    with a  resolution of 5cm by  pixel and 4 channels  are available:
    Red, Blue,  Green and Infrared  (RGBIR).  A Digital  Surface Model
    with  the  same  resolution  is  also  provided  and  a  so-called
    normalized  DSM   representing  the   height  above   ground.  The
    ground-truth  for  24  tiles  are provided  with  6  classes:  Low
    vegetation, High vegetation, Impervious surfaces, Buildings, Cars,
    Clutter.   Three tiles  have  been  used in  this  work, they  are
    displayed             in            Figure~\ref{fig:potsdam-expl}.
    Table~\ref{tab:potsdam} summarizes  the number of samples  of each
    class.

    \begin{figure}[!t]
        \centering
        \begin{tabular}{c@{~}c@{~}c}
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_11_RGB.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_12_RGB.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_3_10_RGB.jpg} \\
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_11_label.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_12_label.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_3_10_label.jpg} \\
            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_05_11_normalized_lastools.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_05_12_normalized_lastools.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_03_10_normalized_lastools.jpg} \\
            {\bfseries{(a)}} & {\bfseries{(b)}}  & {\bfseries{(c)}}\\
        \end{tabular}
        \caption{From top to bottom, true color composition, ground-truth and normalized DSM of: {\bfseries{(a)}} tile 5\_11, {\bfseries{(b)}} tile 5\_12 and {\bfseries{(c)}} tile 3\_10.\label{fig:potsdam-expl}}
    \end{figure}

    \begin{table}[!t]
        \centering
        \caption{Information classes for the three tiles of the Postdam dataset.\label{tab:potsdam}}
        \begin{tabular}[b]{lrrr}\toprule
          & \multicolumn{3}{c}{Number of samples per tile}\\
            \cmidrule{2-4}
            Class &   5\_11  &   5\_12  & 3\_10 \\
          \midrule
          Clutter             & 1,078,611  & 812,038    & 1,890,467 \\
          Trees               & 4,493,295  & 2,132,368  & 8,780,245 \\
          Cars                & 900,076    & 1,101,541  & 434,615 \\
          Buildings           & 13,469,575 & 17,501,421 & 5,128,149 \\
          Low vegetation      & 4,718,219  & 3,210,596  & 11,428,326 \\
          Impervious surfaces & 11,340,224 & 11,242,036 & 8,338,198 \\
          \bottomrule
        \end{tabular}
    \end{table}

    In order to increase the dimensionality of the data, the following features are computed using the RGBIR images similarly to \cite{tuia2015multiclass}:
    \begin{itemize}
        \item 15 Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2~\cite{otb}.
        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features)~\cite{fauvel2013advances};
        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features)~\cite{dalla2010morphological}.
        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)~\cite{dalla2010morphological}.
        \item Textural features for  each channel with neighborhood of
          19x19 pixels:  mean, standard  deviation, range  and entropy
          (16 features)~\cite{otb}.
    \end{itemize}
    The normalized DSM and the raw  RGBIR image are added to these 191
    features and  then then  stacked to  create a  new image  with 196
    bands. %  The  training is  made with  tile 5\_11  and test  with two
    % tiles: 5\_12 which very similar with  tile 5\_11 and 3\_10 which a
    % further area a bit less urban and more industrial.

\section{Experimental results}
\label{sec:test}

    \subsection{Method}
    \label{sec:method}

    The aim  of the experiments is  to compare the proposed  method to
    some   standard  classifiers   used   in   operational  land   map
    production~\cite{rs70912356}. A non-optimized  previous version of
    the method has been already compared to other selection methods in
    \cite{fauvel2015fast}.  Hence, the primary  objective is to assess
    the  operational  efficiency  and  it is  compared  to  other  OTB
    classifiers   used  operationally   through  their   command  line
    applications\footnote{\url{http://otbcb.readthedocs.io/en/latest/OTB-Applications.html}}.

    The following classifiers are tested:
    \begin{itemize}
        \item A k-nearest-neighbors classifier (KNN) with OTB default parameters (32 as number of neighbors).
        \item A Random Forest classifier with parameters optimized by grid search (200 trees, 40 as max depth, 50 as size of the randomly selected subset of features at each tree node)
        \item A GMM classifier with ridge regularization (GMM ridge) with regularization constant optimized by grid search.
    \end{itemize}
    The GMM classifier is part of the external module described in Section~\ref{sec:otb-module}.

    All these classifiers are compared with 3 configurations of the proposed GMM classifier:
    \begin{itemize}
    \item one with forward selection and JM distance as criterion (GMM SFS JM);
    \item one with forward selection and Cohen's kappa as criterion (GMM SFS kappa);
    \item one with floating forward selection and JM distance as criterion (GMM SFFS JM).
    \end{itemize}
    Other configurations  have been  investigated and  performs either
    equally     or     lower     in    terms     of     classification
    accuracy~\cite{al:report}. For the sake of clarity, only the three
    aforementioned configurations are discussed here.

    The training set has been created  with an equal number of samples
    for each class and additionally  a spatial stratification has been
    performed, \emph{i.e.}, each training  sample belongs to a spatial
    polygon  that  does  not  intersect  spatially  with  any  spatial
    polygons used  for the validation.   Several size of  training set
    have been  tested.  For  the Aisa  dataset, experiments  have been
    conducted using  250, 500 and  1000 samples  by class and  for the
    Potsdam dataset, 1000 and 50000 samples by class.

    For SFS and  SFFS selection, the number of variables  to select is
    set to  30 for the  Aisa dataset and  60 for the  Potsdam dataset.
    After  the selection  procedure,  the optimal  number of  selected
    variable is selected  as follow. Rather than  selecting the number
    of variables corresponding to the  highest value of the criterion,
    the number of  retained variables is set when  the criterion stops
    to increase significantly.  It is found by  computing the discrete
    derivative of the criteria. See \ref{fig:crit-evol}.

    \begin{figure}[!t]
        \centering
        \begin{tikzpicture}
            \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa},small]
                \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl.txt};
                \draw[thick,red] (axis cs:15,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:15,\pgfkeysvalueof{/pgfplots/ymax});
                \draw[thick,black] (axis cs:47,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:47,\pgfkeysvalueof{/pgfplots/ymax});
            \end{axis};
            % \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa}]
            %     \addplot[thick,black] table[x=nb,y=crit] {criterion_50000spl.txt};
            %     \draw[thick,red] (axis cs:31,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:31,\pgfkeysvalueof{/pgfplots/ymax});
            %     \draw[thick,black] (axis cs:31,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:31,\pgfkeysvalueof{/pgfplots/ymax});
            % \end{axis};
        \end{tikzpicture}
        \caption{Criterion evolution in function of the number of selected variables. Red vertical line is the retained number of variables and black vertical line is the maximum of the criterion.\label{fig:crit-evol}}
    \end{figure}

    % Then, a small processing is made to select the number of variables used for prediction. The evolution of the criterion function is retrieved from the model file and the discrete derivative of the criterion \emph{DJ} is computed and normalized by its maximum. The subset of variables used for classification $\Omega^{*}$ is the first subset for which this value is lower than a given threshold \emph{th}, equal  to 0.1\% in this experiment,
    % \begin{equation*}
    %     \Omega^{*} = \Omega_{\min \{k|DJ_k<th\}}
    % \end{equation*}
    % with $DJ_k = \frac{J(\Omega_k) - J(\Omega_{k-1})}{\max_k (J(\Omega_k) - J(\Omega_{k-1}))}$.

    The    classification   rate    is    presented   using    Cohen's
    kappa. Processing  time has been  evaluated on a  desktop computer
    with  8Gb of  RAM  and  Intel(R) Core(TM)  i5-3570  CPU @  3.40GHz
    $\times$ 4 processors.

    \subsection{Aisa dataset}
    \label{sec:aisa}

    When creating training and validation  sets, special care is taken
    to assure that training samples are picked out from distinct areas
    than test  samples.  The  polygons of the  reference are  split in
    smaller polygons and then 50\%  of the polygons are taken randomly
    for training and the remaining 50\% for validation.  An example of
    training      and     validation      set     is      shown     in
    Figure~\ref{fig:set-aisa}.  From the  training  polygons, a  given
    number of samples  were selected to build the  training set, while
    all  the pixels  from the  validation polygons  were used  for the
    validation.  Moreover 20  random trials were run  with a different
    training  set (different  polygons).  Table~\ref{tab:aisa-otbsimu}
    presents  the results  of the  experiment with  mean and  standard
    deviation  of  the  Kappa  coefficient  over  the  20  trials  and
    Table~\ref{tab:aisa-otbsimu-time}  the   corresponding  processing
    time.

    \begin{figure}[!t]
        \centering
        \begin{tabular}{c}
            \includegraphics[width=0.7\columnwidth]{Fig/aisa_gt_train.png} \\
            {\bfseries{(a)}} \\
            \includegraphics[width=0.7\columnwidth]{Fig/aisa_gt_test.png} \\
            {\bfseries{(b)}} \\
        \end{tabular}
        \caption{Aisa dataset: {\bfseries{(a)}} training polygons of first trial, {\bfseries{(b)}} test polygons of first trial.\label{fig:set-aisa}}
    \end{figure}

    \begin{table}[!t]
        \centering
        \caption{Average classification accuracy 20 trials (standard deviation in parenthesis).\label{tab:aisa-otbsimu}}
        \begin{tabular}{lccc}\toprule
             & \multicolumn{3}{c}{\bfseries Cohen's kappa} \\ \cmidrule{2-4}
            \# samples by class & 250 & 500 & 1000 \\ \midrule

            GMM SFS kappa & {\bfseries 0.684 (0.025)} & {\bfseries 0.698 (0.027)} & {\bfseries 0.711 (0.028)} \\
            GMM SFS JM &    {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
            GMM SFFS JM &   {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
            GMM ridge &     0.611 (0.040) & 0.620 (0.036) & 0.642 (0.034) \\
            KNN &           0.551 (0.035) & 0.563 (0.033) & 0.574 (0.030) \\
            Random Forest & 0.645 (0.026) & 0.673 (0.023) & {\bfseries 0.693 (0.023)} \\
            \bottomrule
        \end{tabular}
    \end{table}

    \begin{table}[!t]
        \centering
        \caption{Mean processing time for training and classification for results in Table \ref{tab:aisa-otbsimu}.\label{tab:aisa-otbsimu-time}}
        \begin{tabularx}{\columnwidth}{l*{6}{>{\centering\arraybackslash}X}}
            \toprule
             & \multicolumn{3}{c}{\bfseries Training time (s)} & \multicolumn{3}{c}{\bfseries Classification time (s)} \\ \cmidrule{2-7}
            \# samples by class & 250 & 500 & 1000 & 250 & 500 & 1000 \\ \midrule

            GMM SFS kappa & 257             & 496             & 955             & {\bfseries 7.7} & {\bfseries 8.6} & {\bfseries 8.7} \\
            GMM SFS JM &    {\bfseries 8.6} & {\bfseries 8.9} & {\bfseries 9.1} & {\bfseries 9.7} & {\bfseries 9.8} & {\bfseries 9.6} \\
            GMM SFFS JM &   {\bfseries 8.8} & {\bfseries 9.0} & {\bfseries 9.3} & {\bfseries 9.7} & {\bfseries 9.8} & {\bfseries 9.8} \\
            GMM ridge &     71.7            & 105             & 167             & 530 & 530 & 530 \\
            KNN &           {\bfseries 8.9} & 19.6            & 59.7            & 387 & 639 & 887 \\
            Random Forest & 24.5            & 49.3            & 105             & 33.0 & 41.7 & 45.9 \\
            \bottomrule
        \end{tabularx}
    \end{table}

    The results show that, on this  dataset, GMM classifiers with feature selection get the best classification rate. Among the three variations of the selection algorithm, none appears to perform better than the others. Using kappa or Jeffries-Matusita distance as criterion is equal and using SFFS does not give any advantage.

    The difference with the second best classifier, Random Forest, appears to be significant when using 250 and 500 samples. Random Forest has similar performance in term of classification rate with 1000 samples. The GMM classifier with ridge regularization and the KNN classifier are both outperformed.

    In term of computational time, the GMM classifiers are as expected very fast for classification and also for training, except when the criterion function is a classification rate. In this case, using JM distance as criterion and SFS as search strategy is the best choice in term of time efficiency. The good performance in time can be explained by the dimension reduction. Actually, the decision rule corresponding to Equation~(\ref{eq:decision}) has a complexity in $d^3$ where $d$ is the dimension. Thus, reducing $d$ induces a reduction of the classification time.

    The processing times of the three standard classifiers suffer from the increase of training samples. For the GMM classifier with ridge, the selection of regularization parameter is more costly with more samples because of the classification rate estimation needed. For the KNN classifier, the model stores all the training samples and the prediction implies to compute the distance to all the training samples which explains the increase of the processing time and additionally of the size of the model file. Finally, for the Random Forest classifier, the trees tends to be deeper in order to capture the additional information available with more samples and that explains the increase of the processing time.
    % The computation of the kappa seems  to suffer from  a lack of  parallelization   which could be  improved. ->  plus tard dans la conclusion.

    \subsection{Potsdam dataset}

    For the Potsdam  dataset, training samples were  selected from one
    tile (5\_11)  and validation  samples were all  the pixel  of tile
    5\_12  or  3\_10.  Table~\ref{tab:potsdam-otbsimu}  presents  the
    results in terms of classification accuracy and processing time.

    \begin{table*}[!t]
        \centering
        \caption{Kappa coefficient and processing time for 1,000 samples by class and averaged over 5 trials (standard deviation in parenthesis). Processing times are given in second.\label{tab:potsdam-otbsimu}}
        \begin{tabular}{lcccccc}\toprule
            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cmidrule{2-7}
            GMM SFS kappa & 0.694 (0.002)             & {\bfseries 0.669 (0.005)} & {\bfseries 0.533 (0.008)} & 400 & 310  & 13.2 \\
            GMM SFS JM &    0.624 (0.028)             & 0.631 (0.034)             & 0.461 (0.027)             & 2   & 310  & 11 \\
            GMM SFFS JM &   0.624 (0.028)             & 0.631 (0.034)             & 0.461 (0.027)             & 2.6 & 310  & 11 \\
            GMM ridge &     0.632 (0.007)             & 0.592 (0.010)             & 0.433 (0.008)             & 10  & 2000 & all \\
            KNN &           0.637 (0.005)             & 0.607 (0.005)             & 0.478 (0.002)             & 0.7 & 9500 & all \\
            Random Forest & {\bfseries 0.729 (0.004)} & {\bfseries 0.673 (0.005)} & {\bfseries 0.529 (0.009)} & 20  & 840  & all \\
            \bottomrule
        \end{tabular}
    \end{table*}

    \begin{table*}[!t]
        \centering
        \caption{Kappa coefficient and processing time for 50,000 samples by class and averaged over 5 trials (standard deviation in parenthesis). Processing times are given in second.\label{tab:potsdam-otbsimu-big}}
        \begin{tabular}{lcccccc}\toprule
            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cmidrule{2-7}
            GMM SFS kappa & 0.713 (0.001) & 0.684 (0.001) & 0.531 (0.005) & 20000 & 340 & 29 \\
            GMM SFS JM &    0.560 (0.111) & 0.576 (0.104) & 0.435 (0.085) & 6 & 330 & 10 \\
            GMM SFFS JM &   0.560 (0.111) & 0.576 (0.104) & 0.435 (0.085) & 6.6 & 340 & 10 \\
            GMM ridge &     0.641 (0.015) & 0.611 (0.026) & 0.440 (0.015) & 460 & 2000 & all \\
            Random Forest & {\bfseries 0.851 (0.001)} & {\bfseries 0.715 (0.001)} & {\bfseries 0.573 (0.002)} & 2000 & 2000 & all \\
            \bottomrule
        \end{tabular}
    \end{table*}

    With this second dataset, the Random Forest classifier and the GMM classifier with kappa as selection criterion perform the best in terms of classification accuracy. When using 1000 samples per class, no significant difference of classification rate has been observed on test set. But, with 50,000 samples per class, the Random Forest classifier becomes significantly better.

    It can be guessed that the fact that the Gaussian hypothesis is unlikely to be verified explains the difficulty to extract the finer information available in a larger dataset. The JM distance has stressed in Table~\ref{tab:crit} has a strong Gaussian hypothesis and the same hypotheses could also explain the drop of the classification rate of GMM classifiers with JM distance as criterion.

    The KNN classifier and the GMM classifier with ridge regularization are again outperformed even if they get more stable results than GMM classifiers with JM as criterion.

    The number of variables identified by the selection process shows that selection with JM distance as criterion identifies less relevant samples than with kappa as criterion. Moreover, the GMM classifier with kappa manages to get good performance with only 6.7\ of the variables with 1000 samples and 15\% with 50,000 samples.

    In term of processing time, results are similar than with the Aisa dataset. GMM classifiers with selection are very fast for prediction. For example, the GMM classifier with kappa as criterion is 63\% faster than the Random Forest classifier for prediction with 1000 samples and 83\% faster with 50,000 samples.

\section{Conclusion and perspectives}
\label{sec:conclusion}

An algorithm for the classification of high dimension data as hyperspectral image has been presented. The classifier uses GMM model to select iteratively the most relevant features. Several variation of the algorithms has been explored. Experimentations show that allowing backward step in the selection to discard already selected features do not give significant advantages.

Additionally, from the comparison between all the measures used to rank the features, it has been shown that the Jeffries-Matusita distance can be a fast and accurate solution in some case but suffer from some limitations which remain to be determine more precisely. A possible explanation could be that this distance does not perform well when data distribution is not similar to a Gaussian distribution. More investigation is needed to confirm this hypothesis. In any case, it is possible to use directly kappa as criterion function even if it implies a loss of time.

Finally, experiments show that the developed GMM classifier performs, in most cases, at least as best as standard classifiers without features selection in particular Random Forest and even outperforms all of them in term of classification time.

The second achievement of this work is the development of an efficient implementation using GMM properties to derive fast update rules of the model. The resulting code is available as a remote module of the Orfeo toolbox on Github and makes it possible to process huge quantity of high dimension data.

Finally, an improvement could be made to increase the stability of the selected features. For example, with hyperspectral data, a selection of continuous intervals and not band is a possible solution which has already been explored in \cite{serpico2007extraction}.

The python and C++ code are available freely for download: \url{https://github.com/Laadr/FFFS}, \url{https://github.com/Laadr/otbExternalFastFeaturesSelection}.

\appendices
\section{Proof of update rules}
\label{app:proof-update}
    % \subsubsection{Proposition~(\ref{eq:update-inv-back})}
    %     \begin{proof}
    %         \begin{align*}
    %             \mathbf{A} - \alpha \mathbf{v} \mathbf{v}^t
    %             &= (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \\
    %             &~~~- \alpha (- \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}) (- \frac{1}{\alpha} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1})) \\
    %             &= (\boldsymbol{\Sigma}^{(k-1)})^{-1}
    %         \end{align*}
    %     \end{proof}

\begin{proof}[Proof of proposition~(\ref{eq:update-quad})]
    \begin{alignat*}{3}
    (&\mathbf{x}^{(k)})^t && (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)}
       = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x_k \end{array}\right]
                                                            \left[\begin{array}{cc}
                                                                    \mathbf{A}   & \mathbf{v} \\
                                                                    \mathbf{v}^t & \frac{1}{\alpha}
                                                                  \end{array}\right]
                                                                                   \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
     &= &&\left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t \mathbf{A} + x_k \mathbf{v}^t & (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} \end{array}\right]
                                                                                       \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
     &= &&(\mathbf{x}^{(k-1)})^t \mathbf{A} \mathbf{x}^{(k-1)} + x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + (\mathbf{x}^{(k-1)})^t \mathbf{v} x_k + \frac{(x_k)^2}{\alpha} \\
     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
     & &&+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \alpha \mathbf{v} \mathbf{v}^t) \mathbf{x}^{(k-1)}\\
     & &&+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} \mathbf{v}^t \mathbf{x}^{(k-1)} \\
     & &&+ 2 \frac{x_k}{\alpha} \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha^2}) \\
     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} )^2 \\
     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
    \end{alignat*}
\end{proof}

\begin{proof}[Proof of proposition~(\ref{eq:update-log})]
    \begin{align*}
        \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) =
        &\log \left(\left|
            \bigg[\begin{array}{cc}
                \boldsymbol{\Sigma}^{(k-1)} & \mathbf{u}      \\
                \mathbf{u}^t          & \sigma_{kk} \\
            \end{array}\bigg]
        \right|\right) \\
        &\log \left(\left|
            \bigg[\begin{array}{cc}
                \boldsymbol{\Sigma}^{(k-1)} & \mathbf{0}_{k-1,1}     \\
                \mathbf{u}^t          & 1 \\
            \end{array}\bigg]
            \bigg[\begin{array}{cc}
                \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}      \\
                \mathbf{0}_{k-1,1}^t          & \sigma_{kk} - \mathbf{u}^t(\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}\\
            \end{array}\bigg]
        \right|\right) \\
        &\log \left(\left|
            \bigg[\begin{array}{cc}
                \boldsymbol{\Sigma}^{(k-1)} & \mathbf{0}_{k-1,1}     \\
                \mathbf{u}^t          & 1 \\
            \end{array}\bigg]
        \right|\right)
        +\log \left(\left|
            \bigg[\begin{array}{cc}
                \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}      \\
                \mathbf{0}_{k-1,1}^t          & \alpha \\
            \end{array}\bigg]
        \right|\right) \\
        &\log \left( |\boldsymbol{\Sigma}^{(k-1)}| |1| \right) + \log \left( |\mathbf{I}_{k-1}| |\alpha| \right)\\
        &\log \left( |\boldsymbol{\Sigma}^{(k-1)}| \right) + \log ( \alpha )
    \end{align*}
\end{proof}

% \section*{Acknowledgment}

% The authors would like to thank...

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,biblio}

% biography section
%
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

% \begin{IEEEbiography}{Michael Shell}
% Biography text here.
% \end{IEEEbiography}

% % if you will not have a photo at all:
% \begin{IEEEbiographynophoto}{John Doe}
% Biography text here.
% \end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\end{document}

%\begin{figure}[!t]

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: